---
title: "Figures_CritTestingTracing"
author: "Weihsueh Chiu"
date: "7/31/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(coda)
library(bayesplot) 
library(ggplot2)
library(stringr)
library(knitr)
library(ggpubr)
library(dplyr)
library(metR)
library(data.table)
library(usmap)
knitr::opts_chunk$set(echo = TRUE, dpi = 300, fig.height=6, fig.width=6,message = FALSE,warning=FALSE)
functiondir <- "../functions"
modeldir <- "../model"
datezero <- "2019-12-31"
source(file.path(functiondir,"get_reopendata_function.R"))
folder <- "../SEIR.reopen.state.2020.07.22"
preddate <- "2020-07-22"
preddateend <- "2020-09-15"
```

## Load results

```{r load full results, echo=FALSE, include=FALSE}

fips_table <- read.csv(file.path(folder,"FIPS_TABLE.csv"),colClasses=c(
  rep("character",4),rep("numeric",2)
))
rownames(fips_table) <- fips_table$Alpha.code
fips_table$Region <- "South"
fips_table$Region[fips_table$Alpha.code %in% c("CT","ME","MA","NH","RI",
                                               "VT","NJ","NY","PA")] <- "Northeast"
fips_table$Region[fips_table$Alpha.code %in% c("IL","IN","MI","OH","WI","IA",
                                               "KS","MN","MO","NE","ND","SD")] <- "Midwest"
fips_table$Region[fips_table$Alpha.code %in% c("AZ","CO","ID","MT","NV","NM","UT","WY",
                                               "AK","CA","HI","OR","WA")] <- "West"

statesarr <- fips_table$Alpha.code[2:52]
## Read prediction files

allpreddat <- data.frame()
for (statenow in statesarr) {
  csvfile <- list.files(path=file.path(folder,statenow),
                           pattern="*prediction.quantiles.csv",
                         full.names=TRUE)
  preddat <- read.csv(csvfile,as.is=TRUE)
  preddat$state <- statenow
  allpreddat <- rbind(allpreddat,preddat)
}
allpreddat$Output_Var[allpreddat$Output_Var=="N_pos"] <- "positiveIncrease"
allpreddat$Output_Var[allpreddat$Output_Var=="D_pos"] <- "deathIncrease"
allpreddat$Output_Var[allpreddat$Output_Var=="CumPosTest"] <- "positive"
allpreddat$Output_Var[allpreddat$Output_Var=="CumDeath"] <- "death"
allpreddat$Date <- as.Date(allpreddat$Time,origin=as.Date(datezero))

# ## Read data files
# alldat.df <- read.csv(file.path(folder,"DAILYTESTDATA.csv"))
# alldat.df$Date <- as.Date(alldat.df$numDate,origin=as.Date(datezero))
# names(alldat.df)[names(alldat.df)=="variable"] <- "Output_Var"
# names(alldat.df)[names(alldat.df)=="value"] <- "Data"
# 
# allcumdat.df <- read.csv(file.path(folder,"CUMULTESTDATA.csv"))
# allcumdat.df$Date <- as.Date(allcumdat.df$numDate,origin=as.Date(datezero))
# names(allcumdat.df)[names(allcumdat.df)=="variable"] <- "Output_Var"
# names(allcumdat.df)[names(allcumdat.df)=="value"] <- "Data"
# 
# # get reopening data
# reopen.df <- get_reopendata(datadir = "../data")
# minreopen<-aggregate(value~State.Abbr,data=reopen.df,min,na.rm=TRUE)
# rownames(minreopen)<-minreopen$State.Abbr

preddat.df <- subset(allpreddat,Output_Var=="Refft")
preddat.df <- preddat.df[preddat.df$Date==as.Date(preddate),]
preddat.df$Category <- "R[eff](t) ~ 1"
preddat.df$Category[preddat.df$Prediction.75.<1] <- "R[eff](t) < 1"
preddat.df$Category[preddat.df$Prediction.25.>1] <- "R[eff](t) > 1"
preddat.df$Category<-factor(preddat.df$Category,
                            levels=
                              c("R[eff](t) < 1","R[eff](t) ~ 1","R[eff](t) > 1"))
rownames(preddat.df)<- preddat.df$state
```


```{r load single time point scenario results, echo=FALSE, include=FALSE}
folder <- "../SEIR.reopen.state.2020.07.22"
numDate <- as.numeric(as.Date("2020-07-22"))-as.numeric(as.Date(datezero))
t <- numDate
  scenparms <- data.frame()
  for (statenow in statesarr) { 
    sampfile <- list.files(path=file.path(folder,statenow),
                             pattern="*OneTime.out",
                           full.names=TRUE)
    parms <- read.delim(sampfile,as.is=TRUE,sep="\t")
    names(parms)<-gsub("_1.1","",names(parms))
    names(parms)<-gsub("GM_","",names(parms))
    parms$state <- factor(statenow,levels=(sort(statesarr)))
    
    ## Time dependent parameters, evaluated at last day
    parms$ThetaFit <- (parms$ThetaMin - 
                          (parms$ThetaMin - 1)*
                          exp(-((t-60)/parms$TauTheta)^parms$PwrTheta));
    ## Reopening - increase contacts/day
    parms$TimeReopen = 60+parms$TauTheta+parms$TauS;
    parms$ReopenStart = (1 - 1/(1 + exp(4*(t - parms$TimeReopen))));
    parms$ReopenStop =  (1 - 1/(1 + exp(4*(t - (parms$TimeReopen+parms$TauR)))));
    parms$ReopenPct = ((t - parms$TimeReopen)*(parms$rMax/parms$TauR)*(parms$ReopenStart-parms$ReopenStop)+parms$rMax*parms$ReopenStop);
    ## Contacts/day
    parms$c = parms$c0 * (parms$ThetaFit + (1 - parms$ThetaMin) * parms$ReopenPct); 
    ## Hygiene - reduce infection probability/infected contact
    parms$HygieneFit <- parms$ThetaFit^parms$HygienePwr;
    parms$beta <- parms$beta0 * parms$HygieneFit; # infection probability/infected contact
    parms$Delta <- (parms$c*parms$beta -
                              parms$c0*parms$beta0*
                              parms$ThetaMin^(1+parms$HygienePwr))/
    (parms$c0*parms$beta0 -
       parms$c0*parms$beta0*
       parms$ThetaMin^(1+parms$HygienePwr))
    ## Time dependence of testing/contact tracting
    parms$TestingTimeDep <- (1-1/(1+exp((t-parms$T50Testing)/parms$TauTesting))); 
    ## Contact tracing
    parms$FTraced <- parms$FracTraced * parms$TestingTimeDep;
    ## Testing
    parms$lambda <- parms$TestingTimeDep * parms$lambda0; 
    parms$lambda_C <- parms$TestingTimeDep * parms$lambda0_C; 
    parms$rho_C <- parms$TestingTimeDep * parms$rho0_C;
    parms$fracpos = parms$FTraced*parms$lambda_C/(parms$lambda_C + parms$rho_C)+(1-parms$FTraced)*parms$lambda/(parms$lambda+parms$rho); # fraction of infected that are tested and positive
    ## Case fatality
    parms$fracposmin <- parms$IFR / 0.9; # max 90% of cases fatal
    parms$CFR <- ifelse(parms$fracpos > parms$fracposmin, parms$IFR/parms$fracpos, 0.9); # Adjust infected fatality to (tested) case fatality
    parms$delta <- parms$rho * parms$CFR/(1-parms$CFR);
    scenparms <- rbind(scenparms,parms)
}
scenparms<-scenparms[,!duplicated(names(scenparms))]
scenparms.med<-aggregate(.~state,data=scenparms,median)
scenparms$Category <- preddat.df[as.character(scenparms$state),"Category"]

```

## Critical values of testing and tracing rates to achieve R[eff](t)<1

If critical value of testing > 1 per day, then an infinite value is assigned to represent that it cannot be achieved.

```{r lambda fc crit,fig.height=6,fig.width=10, echo=FALSE}

fCcritfunc<-function(scenparms) {
  fCcrit <- (1 - 1/(scenparms$S*scenparms$c*scenparms$beta*
                       ((1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
                          scenparms$FAsymp/scenparms$rho)))
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - fCcrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  #fCcrit[scenparms$Refft<=1]<-scenparms$FTraced[scenparms$Refft<=1]
  fCcrit[fCcrit<0]<-0
  return(fCcrit)
}

lambdacritfunc<-function(scenparms) {
  lambdacrit <- scenparms$rho*
    (scenparms$S*scenparms$c*scenparms$beta*(1-scenparms$FTraced) - scenparms$rho)/
    (scenparms$rho - scenparms$FAsymp*scenparms$S*scenparms$c*scenparms$beta*(1-scenparms$FTraced))
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+lambdacrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  lambdacrit[lambdacrit > 1] <- Inf
  #lambdacrit[scenparms$Refft<=1]<-scenparms$lambda[scenparms$Refft<=1]
  lambdacrit[lambdacrit<0]<-0
  return(lambdacrit)
}

mubothcritfunc<-function(scenparms) {
  tmp_b <- (scenparms$rho+
              scenparms$S*scenparms$c*scenparms$beta*(
                scenparms$rho - scenparms$FAsymp
              ))/
    (scenparms$S*scenparms$c*scenparms$beta*scenparms$FAsymp)
  tmp_c <- scenparms$rho*(scenparms$rho - scenparms$S*scenparms$c*scenparms$beta)/
    (scenparms$S*scenparms$c*scenparms$beta*scenparms$FAsymp)
  mubothcrit <- (-tmp_b + sqrt(tmp_b^2 - 4*tmp_c))/2
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - mubothcrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+mubothcrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  mubothcrit[mubothcrit > 1] <- Inf
  #mubothcrit[scenparms$Refft<=1] <- NA
  return(mubothcrit)
}

tmpdf <- cbind(scenparms[,c("state","Rt","Refft","Category","lambda","FTraced")],
               data.table(fCcrit=fCcritfunc(scenparms),
                          lambdacrit=lambdacritfunc(scenparms),
                          mubothcrit=mubothcritfunc(scenparms))
                          )

tmpdf.quantiles<-aggregate(.~state,data=tmpdf[,c("state","fCcrit","lambdacrit","mubothcrit")],
                           quantile,prob=c(0.025,0.05,0.25,0.5,0.75,0.95,0.975))
fwrite(as.data.table(tmpdf.quantiles),"TestTraceCritValues.csv")


tmpdf.df<-melt(as.data.table(tmpdf),id.vars=c("state","Category","Rt","Refft","lambda","FTraced"),
            measure.vars = c("fCcrit","lambdacrit","mubothcrit"))
tmpdf.df$variable<-factor(tmpdf.df$variable,levels=c(
  "lambdacrit","fCcrit","mubothcrit"
))
#tmpdf.df$value[is.infinite(tmpdf.df$value)]<-1e6

current.df <- aggregate(.~state,data=tmpdf[,c("state","Category","lambda","FTraced")],median)
tmpdf2.df <- data.table(state=rep(current.df$state,3),
                        variable=c(rep("lambdacrit",nrow(current.df)),
                                   rep("fCcrit",nrow(current.df)),
                                   rep("mubothcrit",nrow(current.df))),
                        lambda=rep(current.df$lambda,3),
                        FTraced=rep(current.df$FTraced),3)
tmpdf2.df$variable<-factor(tmpdf2.df$variable,levels=c(
  "lambdacrit","fCcrit","mubothcrit"
))
tmpdf2.df[tmpdf2.df$variable=="lambdacrit","FTraced"]<-NA
tmpdf2.df[tmpdf2.df$variable=="fCcrit","lambda"]<-NA

dflabs<-c("Critical Testing Rate","Critical Tracing Fraction",
          "Critical Combined Testing Rate and Tracing Fraction")
names(dflabs)<-c("lambdacrit","fCcrit","mubothcrit")

ptesttrace<-ggplot(data=tmpdf.df,
       aes(x=state,
           y=value)) + 
    geom_boxplot(aes(fill=Category),outlier.shape = NA,position="dodge")+
  geom_point(aes(x=state,y=lambda,shape="Current Testing Rate (Median estimate)"),data=tmpdf2.df)+
  geom_point(aes(x=state,y=FTraced,shape="Current Fraction Traced (Median estimate)"),data=tmpdf2.df)+
  scale_fill_viridis_d(begin=0.9,end=0.5,option="magma",
                       labels=list(bquote(R[eff](t)<1),
                                   bquote(R[eff](t)%~~%1),
                                   bquote(R[eff](t)>1)))+
  #ylab(bquote(paste("Percent increase in both testing and tracing for ",R[eff](t)<1)))+
  ylab(bquote(paste("Testing rate ",(d^{-1})," and/or fraction traced for ",R[eff](t)<1)))+
  labs(fill="",color="",shape="")+
  theme_bw()+theme(legend.position="bottom")+
  geom_hline(yintercept=0)+
  facet_wrap(~variable,ncol=1,
             labeller = labeller(variable=dflabs))
print(ptesttrace)
ggsave("FigS11_TestingTracing.2020.07.22.pdf",plot=ptesttrace,height=4,width=6,scale=1.75)
ggsave("FigS11_TestingTracing.2020.07.22.jpeg",plot=ptesttrace,height=4,width=6,scale=1.75)

```


## Critical values of testing and tracing rates to achieve R(t)<1

Critical testing/tracing using R(t) and assuming complete reopening (using c0 and beta0 instead of c and beta). If critical value of testing > 1 per day, then an infinite value is assigned to represent that it cannot be achieved.fitparms.df$variable[fitparms.df$variable=="thetamin"]<-"ThetaMin"


```{r mu crit using Rt,fig.height=6,fig.width=10, echo=FALSE}

fCreopencritfunc<-function(scenparms) {
  fCreopencrit <- (1 - 1/(scenparms$c*scenparms$beta*
                       ((1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
                          scenparms$FAsymp/scenparms$rho)))
  # checkRt <- scenparms$c*scenparms$beta*(1 - fCreopencrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRt,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  #fCreopencrit[scenparms$Rt<=1]<-scenparms$FTraced[scenparms$Rt<=1]
  fCreopencrit[fCreopencrit<0]<-0
  return(fCreopencrit)
}

lambdareopencritfunc<-function(scenparms) {
  lambdareopencrit <- scenparms$rho*
    (scenparms$c*scenparms$beta*(1-scenparms$FTraced) - scenparms$rho)/
    (scenparms$rho - scenparms$FAsymp*scenparms$c*scenparms$beta*(1-scenparms$FTraced))
  # checkRt <- scenparms$S*scenparms$c*scenparms$beta*(1 - scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+lambdareopencrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRt,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  lambdareopencrit[lambdareopencrit > 1] <- Inf
  #lambdareopencrit[scenparms$Rt<=1]<-scenparms$lambda[scenparms$Rt<=1]
  lambdareopencrit[lambdareopencrit<0]<-0
  return(lambdareopencrit)
}

mubothreopencritfunc<-function(scenparms) {
  tmp_b <- (scenparms$rho+
              scenparms$c*scenparms$beta*(
                scenparms$rho - scenparms$FAsymp
              ))/
    (scenparms$c*scenparms$beta*scenparms$FAsymp)
  tmp_c <- scenparms$rho*(scenparms$rho - scenparms$c*scenparms$beta)/
    (scenparms$c*scenparms$beta*scenparms$FAsymp)
  mubothreopencrit <- (-tmp_b + sqrt(tmp_b^2 - 4*tmp_c))/2
  # checkRt <- scenparms$c*scenparms$beta*(1 - mubothreopencrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+mubothreopencrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRt,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  mubothreopencrit[mubothreopencrit > 1] <- Inf
  #mubothreopencrit[scenparms$Rt<=1] <- NA
  return(mubothreopencrit)
}

tmpdf <- cbind(scenparms[,c("state","Rt","Refft","Category","lambda","FTraced")],
               data.table(fCreopencrit=fCreopencritfunc(scenparms),
                          lambdareopencrit=lambdareopencritfunc(scenparms),
                          mubothreopencrit=mubothreopencritfunc(scenparms))
                          )

tmpdf.quantiles<-aggregate(.~state,data=tmpdf[,c("state","fCreopencrit","lambdareopencrit","mubothreopencrit")],
                           quantile,prob=c(0.025,0.05,0.25,0.5,0.75,0.95,0.975))
fwrite(as.data.table(tmpdf.quantiles),"RtReopen_TestTraceCritValues.csv")



tmpdf.df<-melt(as.data.table(tmpdf),id.vars=c("state","Category","Rt","Refft","lambda","FTraced"),
            measure.vars = c("fCreopencrit","lambdareopencrit","mubothreopencrit"))
tmpdf.df$variable<-factor(tmpdf.df$variable,levels=c(
  "lambdareopencrit","fCreopencrit","mubothreopencrit"
))
#tmpdf.df$value[is.infinite(tmpdf.df$value)]<-1e6

current.df <- aggregate(.~state,data=tmpdf[,c("state","Category","lambda","FTraced")],median)
tmpdf2.df <- data.table(state=rep(current.df$state,3),
                        variable=c(rep("lambdareopencrit",nrow(current.df)),
                                   rep("fCreopencrit",nrow(current.df)),
                                   rep("mubothreopencrit",nrow(current.df))),
                        lambda=rep(current.df$lambda,3),
                        FTraced=rep(current.df$FTraced),3)
tmpdf2.df$variable<-factor(tmpdf2.df$variable,levels=c(
  "lambdareopencrit","fCreopencrit","mubothreopencrit"
))
tmpdf2.df[tmpdf2.df$variable=="lambdareopencrit","FTraced"]<-NA
tmpdf2.df[tmpdf2.df$variable=="fCreopencrit","lambda"]<-NA

dflabs<-c("Critical Testing Rate","Critical Tracing Fraction",
          "Critical Combined Testing Rate and Tracing Fraction")
names(dflabs)<-c("lambdareopencrit","fCreopencrit","mubothreopencrit")

preopentesttrace<-ggplot(data=tmpdf.df,
       aes(x=state,
           y=value)) + 
    geom_boxplot(aes(fill=Category),outlier.shape = NA,position="dodge")+
  geom_point(aes(x=state,y=lambda,shape="Current Testing Rate (Median estimate)"),data=tmpdf2.df)+
  geom_point(aes(x=state,y=FTraced,shape="Current Fraction Traced (Median estimate)"),data=tmpdf2.df)+
  scale_fill_viridis_d(begin=0.9,end=0.5,option="magma",
                       labels=list(bquote(R[eff](t)<1),
                                   bquote(R[eff](t)%~~%1),
                                   bquote(R[eff](t)>1)))+
  #ylab(bquote(paste("Percent increase in both testing and tracing for ",R(t)<1)))+
  ylab(bquote(paste("Testing rate ",(d^{-1})," and/or fraction traced for ",R(t)<1)))+
  labs(fill="",color="",shape="")+
  theme_bw()+theme(legend.position="bottom")+
  geom_hline(yintercept=0)+
  facet_wrap(~variable,ncol=1,
             labeller = labeller(variable=dflabs))
print(preopentesttrace)

ggsave("FigS12_TestingTracingReopen.2020.07.22.pdf",plot=preopentesttrace,height=4,width=6,scale=1.75)
ggsave("FigS12_TestingTracingReopen.2020.07.22.jpeg",plot=preopentesttrace,height=4,width=6,scale=1.75)

```


## Critical values of fold-increases in testing and tracing to achieve R[eff](t)<1

If critical value of testing > 1 per day, then an infinite value is assigned to represent that it cannot be achieved.

```{r mu crit,fig.height=6,fig.width=10, echo=FALSE}

muCcritfunc<-function(scenparms) {
  muCcrit <- (1 - 1/(scenparms$S*scenparms$c*scenparms$beta*
                       ((1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
                          scenparms$FAsymp/scenparms$rho)))/scenparms$FTraced
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - muCcrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  muCcrit[scenparms$Refft<=1]<-1
  muCcrit[muCcrit<1]<-1
  return(muCcrit)
}

mulambdacritfunc<-function(scenparms) {
  mulambdacrit <- (scenparms$rho/scenparms$lambda)*
    (scenparms$S*scenparms$c*scenparms$beta*(1-scenparms$FTraced) - scenparms$rho)/
    (scenparms$rho - scenparms$FAsymp*scenparms$S*scenparms$c*scenparms$beta*(1-scenparms$FTraced))
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+mulambdacrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  mulambdacrit[mulambdacrit*scenparms$lambda > 1] <- Inf
  mulambdacrit[scenparms$Refft<=1]<-1
  mulambdacrit[mulambdacrit<1]<-1
  return(mulambdacrit)
}

mubothcritfunc<-function(scenparms) {
  tmp_b <- (scenparms$rho*scenparms$lambda+
              scenparms$S*scenparms$c*scenparms$beta*(
                scenparms$FTraced*scenparms$rho - scenparms$FAsymp*scenparms$lambda
              ))/
    (scenparms$S*scenparms$c*scenparms$beta*scenparms$lambda*scenparms$FTraced*scenparms$FAsymp)
  tmp_c <- scenparms$rho*(scenparms$rho - scenparms$S*scenparms$c*scenparms$beta)/
    (scenparms$S*scenparms$c*scenparms$beta*scenparms$lambda*scenparms$FTraced*scenparms$FAsymp)
  mubothcrit <- (-tmp_b + sqrt(tmp_b^2 - 4*tmp_c))/2
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - mubothcrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+mubothcrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  mubothcrit[mubothcrit*scenparms$lambda > 1] <- Inf
  mubothcrit[scenparms$Refft<=1]<-1
  mubothcrit[mubothcrit<1]<-1
  return(mubothcrit)
}

tmpdf <- cbind(scenparms[,c("state","Rt","Refft","Category")],
               data.table(muCcrit=muCcritfunc(scenparms),
                          mulambdacrit=mulambdacritfunc(scenparms),
                          mubothcrit=mubothcritfunc(scenparms)))

tmpdf.quantiles<-aggregate(.~state,data=tmpdf[,c("state","muCcrit","mulambdacrit","mubothcrit")],
                           quantile,prob=c(0.025,0.05,0.25,0.5,0.75,0.95,0.975))
fwrite(as.data.table(tmpdf.quantiles),"TestTraceCritValues.csv")


tmpdf.df<-melt(as.data.table(tmpdf),id.vars=c("state","Category","Rt","Refft"),
            measure.vars = c("muCcrit","mulambdacrit","mubothcrit"))
tmpdf.df$variable<-factor(tmpdf.df$variable,levels=c(
  "mulambdacrit","muCcrit","mubothcrit"
))
tmpdf.df$value[is.infinite(tmpdf.df$value)]<-1e6

dflabs<-c("Increase Testing","Increase Tracing","Increase Testing and Tracing")
names(dflabs)<-c("mulambdacrit","muCcrit","mubothcrit")

ptesttrace<-ggplot(data=tmpdf.df,
       aes(x=state,
           y=value,fill=Category
           )) + 
    geom_boxplot(outlier.shape = NA,position="dodge")+
  scale_fill_viridis_d(begin=0.9,end=0.5,option="magma",
                       labels=list(bquote(R[eff](t)<1),
                                   bquote(R[eff](t)%~~%1),
                                   bquote(R[eff](t)>1)))+
  #ylab(bquote(paste("Percent increase in both testing and tracing for ",R[eff](t)<1)))+
  ylab(bquote(paste("Fold increase in testing and/or tracing for ",R[eff](t)<1)))+
  scale_y_continuous( labels = function(x) paste0(x, "X"),breaks=1:6)+
  labs(fill="",color="",shape="")+
  theme_bw()+theme(legend.position="bottom")+
  geom_hline(yintercept=0)+
  coord_cartesian(ylim=c(1,6))+
  facet_wrap(~variable,ncol=1,
             labeller = labeller(variable=dflabs))
print(ptesttrace)
ggsave("FigS11_TestingTracing.2020.07.22.pdf",plot=ptesttrace,height=4,width=6,scale=1.75)
ggsave("FigS11_TestingTracing.2020.07.22.jpeg",plot=ptesttrace,height=4,width=6,scale=1.75)

```


## Critical values of fold-increases in testing and tracing to achieve R(t)<1

Critical testing/tracing using R(t) and assuming complete reopening (using c0 and beta0 instead of c and beta). If critical value of testing > 1 per day, then an infinite value is assigned to represent that it cannot be achieved.fitparms.df$variable[fitparms.df$variable=="thetamin"]<-"ThetaMin"


```{r mu crit using Rt,fig.height=6,fig.width=10, echo=FALSE}
muCcritreopenfunc<-function(scenparms) {
  muCcrit <- (1 - 1/(scenparms$c0*scenparms$beta0*
                       ((1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
                          scenparms$FAsymp/scenparms$rho)))/scenparms$FTraced
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - muCcrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  muCcrit[muCcrit<1]<-1
  return(muCcrit)
}

mulambdacritreopenfunc<-function(scenparms) {
  mulambdacrit <- (scenparms$rho/scenparms$lambda)*
    (scenparms$c0*scenparms$beta0*(1-scenparms$FTraced) - scenparms$rho)/
    (scenparms$rho - scenparms$FAsymp*scenparms$S*scenparms$c*scenparms$beta*(1-scenparms$FTraced))
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+mulambdacrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  mulambdacrit[mulambdacrit*scenparms$lambda > 1] <- Inf
  mulambdacrit[mulambdacrit<1]<-1
  return(mulambdacrit)
}

mubothcritreopenfunc<-function(scenparms) {
  tmp_b <- (scenparms$rho*scenparms$lambda+
              scenparms$c0*scenparms$beta0*(
                scenparms$FTraced*scenparms$rho - scenparms$FAsymp*scenparms$lambda
              ))/
    (scenparms$c0*scenparms$beta0*scenparms$lambda*scenparms$FTraced*scenparms$FAsymp)
  tmp_c <- scenparms$rho*(scenparms$rho - scenparms$c0*scenparms$beta0)/
    (scenparms$c0*scenparms$beta0*scenparms$lambda*scenparms$FTraced*scenparms$FAsymp)
  mubothcrit <- (-tmp_b + sqrt(tmp_b^2 - 4*tmp_c))/2
  # checkRefft <- scenparms$S*scenparms$c*scenparms$beta*(1 - mubothcrit*scenparms$FTraced)*(
  #   (1-scenparms$FAsymp)/(scenparms$rho+mubothcrit*scenparms$lambda)+
  #     scenparms$FAsymp/scenparms$rho
  # )
  # print(quantile(checkRefft,prob=c(0,0.01,0.05,0.5,0.95,0.99,1)))
  mubothcrit[mubothcrit*scenparms$lambda > 1] <- Inf
  mubothcrit[mubothcrit<1]<-1
  return(mubothcrit)
}

tmpdf <- cbind(scenparms[,c("state","Rt","Refft","Category")],
               data.table(muCcrit=muCcritreopenfunc(scenparms),
                          mulambdacrit=mulambdacritreopenfunc(scenparms),
                          mubothcrit=mubothcritreopenfunc(scenparms)))


tmpdf.quantiles<-aggregate(.~state,data=tmpdf[,c("state","muCcrit","mulambdacrit","mubothcrit")],
                           quantile,prob=c(0.025,0.05,0.25,0.5,0.75,0.95,0.975))

tmpdf.df<-melt(as.data.table(tmpdf),id.vars=c("state","Category","Rt","Refft"),
            measure.vars = c("muCcrit","mulambdacrit","mubothcrit"))
tmpdf.df$variable<-factor(tmpdf.df$variable,levels=c(
  "mulambdacrit","muCcrit","mubothcrit"
))
tmpdf.df$value[is.infinite(tmpdf.df$value)]<-1e6
dflabs<-c("Increase Testing","Increase Tracing","Increase Testing and Tracing")
names(dflabs)<-c("mulambdacrit","muCcrit","mubothcrit")

ptesttracereopen<-ggplot(data=tmpdf.df,
       aes(x=state,
           y=value,fill=Category
           )) + 
    geom_boxplot(outlier.shape = NA,position="dodge")+
  scale_fill_viridis_d(begin=0.9,end=0.5,option="magma",
                       labels=list(bquote(R[eff](t)<1),
                                   bquote(R[eff](t)%~~%1),
                                   bquote(R[eff](t)>1)))+
  #ylab(bquote(paste("Percent increase in both testing and tracing for ",R[eff](t)<1)))+
  ylab(bquote(paste("Fold increase in testing and/or tracing for R(reopen)<1")))+
  scale_y_continuous( labels = function(x) paste0(x, "X"),breaks=1:10)+
  labs(fill="",color="",shape="")+
  theme_bw()+theme(legend.position="bottom")+
  geom_hline(yintercept=0)+
  #scale_y_log10()+
  #ylim(1,10)+
  coord_cartesian(ylim=c(1,10))+
  facet_wrap(~variable,ncol=1,
             labeller = labeller(variable=dflabs))
print(ptesttracereopen)
fwrite(as.data.table(tmpdf.quantiles),"RtReopen_TestTraceCritValues.csv")
# ggsave("FigS12_TestingTracingReopen.2020.07.22.pdf",plot=ptesttracereopen,height=4,width=6,scale=1.75)
# ggsave("FigS12_TestingTracingReopen.2020.07.22.jpeg",plot=ptesttracereopen,height=4,width=6,scale=1.75)


```
