---
title: "Validation Figures"
author: "Weihsueh Chiu"
date: "`r format(Sys.time(), '%F')`"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(coda)
library(bayesplot) 
library(ggplot2)
library(stringr)
library(knitr)
library(ggpubr)
library(dplyr)
library(metR)
library(data.table)
library(usmap)
knitr::opts_chunk$set(echo = TRUE, dpi = 300, fig.height=6, fig.width=6,message = FALSE,warning=FALSE)
functiondir <- "../functions"
modeldir <- "../model"
datezero <- "2019-12-31"
source(file.path(functiondir,"get_reopendata_function.R"))
```

## Validation 

Calibration using data until April 30. Validation with data through June 20.

```{r load validation results, echo=FALSE, include=FALSE}

# Separately source "run_plot_validation_scenario.R")

folder <- "../SEIR.reopen.2020.04.30"
fips_table <- read.csv(file.path(folder,"FIPS_TABLE.csv"),colClasses=c(
  rep("character",4),rep("numeric",2)
))
rownames(fips_table) <- fips_table$Alpha.code
fips_table$Region <- "South"
fips_table$Region[fips_table$Alpha.code %in% c("CT","ME","MA","NH","RI",
                                               "VT","NJ","NY","PA")] <- "Northeast"
fips_table$Region[fips_table$Alpha.code %in% c("IL","IN","MI","OH","WI","IA",
                                               "KS","MN","MO","NE","ND","SD")] <- "Midwest"
fips_table$Region[fips_table$Alpha.code %in% c("AZ","CO","ID","MT","NV","NM","UT","WY",
                                               "AK","CA","HI","OR","WA")] <- "West"

statesarr <- fips_table$Alpha.code[2:52]
## Read prediction files
allcalvaldat <- data.frame()
for (statenow in statesarr) { 
  csvfile <- list.files(path=file.path(folder,statenow),
                           pattern="*prediction.quantiles.csv",
                         full.names=TRUE)
  calvaldat <- read.csv(csvfile,as.is=TRUE)
  calvaldat$state <- statenow
  allcalvaldat <- rbind(allcalvaldat,calvaldat)
}
allcalvaldat$Output_Var[allcalvaldat$Output_Var=="N_pos"] <- "positiveIncrease"
allcalvaldat$Output_Var[allcalvaldat$Output_Var=="D_pos"] <- "deathIncrease"
allcalvaldat$Output_Var[allcalvaldat$Output_Var=="CumPos"] <- "positive"
allcalvaldat$Output_Var[allcalvaldat$Output_Var=="D_T"] <- "death"
allcalvaldat$Date <- as.Date(allcalvaldat$Time,origin=as.Date(datezero))

## Read data files
alldat.df <- read.csv(file.path(folder,"DAILYTESTDATA.csv"))
alldat.df$Date <- as.Date(alldat.df$numDate,origin=as.Date(datezero))
names(alldat.df)[names(alldat.df)=="variable"] <- "Output_Var"
names(alldat.df)[names(alldat.df)=="value"] <- "Data"

allcumdat.df <- read.csv(file.path(folder,"CUMULTESTDATA.csv"))
allcumdat.df$Date <- as.Date(allcumdat.df$numDate,origin=as.Date(datezero))
names(allcumdat.df)[names(allcumdat.df)=="variable"] <- "Output_Var"
names(allcumdat.df)[names(allcumdat.df)=="value"] <- "Data"

```

Example states: one from each major region of the US: New York (Northeast), Ohio (Midwest), Texas (South), and Washington (West).

```{r validation example states, echo=FALSE, fig.height=4.87,fig.width=8}
egstates <- c("OH","NY","TX","WA")

calvaldat.df <- subset(allcalvaldat,state %in% egstates)

## Observed/predicted daily
obsdat.df<-subset(alldat.df,state %in% egstates & (
  Output_Var == "positiveIncrease" | Output_Var == "deathIncrease"
)) 

## Calibration vs. validation
datadatemax <- "2020-04-30"
novalid <- FALSE
validdate <- "2020-06-20"
obsdat.df$Calibration <- obsdat.df$Date <= datadatemax
calvaldat.df$Calibration <- calvaldat.df$Date <= datadatemax
obsdat.df$Calibration <- factor(ifelse(obsdat.df$Calibration,
                                       "Train","Validate"),
                                levels=c("Train","Validate"))

calvaldat.df$Calibration <- factor(ifelse(calvaldat.df$Calibration,
                                        "Train","Validate"),
                                levels=c("Train","Validate"))

if (novalid) {
  obsdat.df <- subset(obsdat.df,Date <= as.Date(datadatemax))
  calvaldat.df <- subset(calvaldat.df,Date <= as.Date(datadatemax))
} else {
  obsdat.df <- subset(obsdat.df,Date <= as.Date(validdate))
  calvaldat.df <- subset(calvaldat.df,Date <= as.Date(validdate))
}
obsnames <- c("Daily reported cases","Daily confirmed deaths")
names(obsnames) <- c("positiveIncrease","deathIncrease")

calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>1000 &
                                calvaldat.df$state == "NY" & calvaldat.df$Output_Var=="deathIncrease"]<-1000
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>200 &
                                calvaldat.df$state == "OH" & calvaldat.df$Output_Var=="deathIncrease"]<-200
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>100 &
                                calvaldat.df$state == "TX" & calvaldat.df$Output_Var=="deathIncrease"]<-100
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>100 &
                                calvaldat.df$state == "WA" & calvaldat.df$Output_Var=="deathIncrease"]<-100
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>12500 &
                                calvaldat.df$state == "NY" & calvaldat.df$Output_Var=="positiveIncrease"]<-12500
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>3000 &
                                calvaldat.df$state == "OH" & calvaldat.df$Output_Var=="positiveIncrease"]<-3000
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>5000 &
                                calvaldat.df$state == "TX" & calvaldat.df$Output_Var=="positiveIncrease"]<-5000
calvaldat.df$Prediction.97.5.[calvaldat.df$Prediction.97.5.>1000 &
                                calvaldat.df$state == "WA" & calvaldat.df$Output_Var=="positiveIncrease"]<-1000


calvaldat.df$Prediction.2.5.<-100000*calvaldat.df$Prediction.2.5./fips_table[
  as.character(calvaldat.df$state),"pop"]
calvaldat.df$Prediction.25.<-100000*calvaldat.df$Prediction.25./fips_table[
  as.character(calvaldat.df$state),"pop"]
calvaldat.df$Prediction.50.<-100000*calvaldat.df$Prediction.50./fips_table[
  as.character(calvaldat.df$state),"pop"]
calvaldat.df$Prediction.75.<-100000*calvaldat.df$Prediction.75./fips_table[
  as.character(calvaldat.df$state),"pop"]
calvaldat.df$Prediction.97.5.<-100000*calvaldat.df$Prediction.97.5./fips_table[
  as.character(calvaldat.df$state),"pop"]

obsdat.df$Data <- 100000*obsdat.df$Data/fips_table[as.character(obsdat.df$state),"pop"]

dummydf <- data.table(state=egstates)
dummydf$Output_Var<-"deathIncrease"
dummydf$Data<-10
dummydf2 <- data.table(state=egstates)
dummydf2$Output_Var<-"positiveIncrease"
dummydf2$Data<-118
dummydf <- rbind(dummydf,dummydf2)
dummydf$Date <- as.Date(validdate)

p<-ggplot(subset(calvaldat.df,Output_Var=="positiveIncrease" | Output_Var=="deathIncrease"))+
      #geom_point(data=dummydf,aes(x=Date,y=Data),shape=NA)+
  geom_col(aes(x=Date,y=Data,fill="",
                             alpha=Calibration),data=obsdat.df)+
  geom_ribbon(aes(x=Date,ymin=Prediction.2.5.,ymax=Prediction.97.5.,
                  fill="CrI",alpha=Calibration))+
  geom_ribbon(aes(x=Date,ymin=Prediction.25.,ymax=Prediction.75.,
                  fill="IQR",alpha=Calibration))+
  geom_line(aes(x=Date,y=Prediction.50.,linetype="Median",
                             alpha=Calibration))+
  scale_fill_viridis_d(option="magma",begin = 0.2,end=0.95) +
  scale_x_date(date_minor_breaks = "1 day")+
  facet_wrap(~Output_Var*state,ncol=4,
                  labeller = labeller(Output_Var = obsnames),
                scales="free_y")+
     # facet_grid(state~Output_Var,
     #              labeller = labeller(Output_Var = obsnames),
     #            scales="free_y")+
  theme_bw()+theme(legend.position = "bottom",
                   legend.title = element_text(size = 8),
                    legend.text = element_text(size = 8),
                   strip.text = element_text(size=10,
                                             margin = 
                                               margin(.05, 0, .05, 0, "cm"))) + 
        scale_alpha_discrete(range=rev(c(0.5,0.8)))+
              geom_vline(xintercept=as.Date(datadatemax)+1,
                         linetype="dotted",color="black") +
  #scale_y_log10(breaks=10^(0:10))+
        #annotation_logticks(sides="l")+
       ylab("Daily reported cases or deaths per 100,000")+
        labs(fill="",linetype="Prediction",alpha="Dataset")+
        guides(alpha = guide_legend(order = 1,nrow=1),
         linetype = guide_legend(order = 2),
         fill = guide_legend(order = 3,nrow=1))


print(p)
ggsave("Fig1B.pdf",plot=p,height=4.5,width=13,scale=0.8)
# ggsave("Fig1B.png",plot=p,height=4.5,width=13,scale=1.1)
# ggsave("Fig1B.jpeg",plot=p,height=4.5,width=13,scale=1.1)

```



Next are scatter plots showing overall calibration and validation of model.

```{r validation scatter plot, echo=FALSE, fig.height=6, fig.width=8.5, fig.cap="Figure S-1. Validation runs: Comparison of data and predictions overall."}
calvaldat.df <- allcalvaldat

## Observed/predicted daily
obsdat.df<-subset(alldat.df,
  Output_Var == "positiveIncrease" | Output_Var == "deathIncrease"
)

## Calibration vs. validation
datadatemax <- "2020-04-30"
novalid <- FALSE
validdate <- "2020-06-20"
obsdat.df$Calibration <- obsdat.df$Date <= datadatemax
calvaldat.df$Calibration <- calvaldat.df$Date <= datadatemax
if (novalid) {
  obsdat.df <- subset(obsdat.df,Date <= as.Date(datadatemax))
  calvaldat.df <- subset(calvaldat.df,Date <= as.Date(datadatemax))
} else {
  obsdat.df <- subset(obsdat.df,Date <= as.Date(validdate))
  calvaldat.df <- subset(calvaldat.df,Date <= as.Date(validdate))
}
obsnames <- c("Daily reported cases","Daily confirmed deaths")
names(obsnames) <- c("positiveIncrease","deathIncrease")

predobs.df <- merge(obsdat.df,calvaldat.df,all.x=TRUE)
predobs.df$DataSet <- ifelse(predobs.df$Calibration,
                             "Calibration","Validation")

likeCIdeath <- data.frame(Data=c(0.1,0.3,seq(1,
                                   max(subset(predobs.df,
                                              Output_Var=="deathIncrease"
                                              )$Data)),1000,3000,10000,20000))
likeCIdeath$Prediction.50. <- likeCIdeath$Data
likeCIdeath$Prediction.2.5. <- qnbinom(0.025,size=8,
                                       prob=8/(8+likeCIdeath$Data))
likeCIdeath$Prediction.97.5. <- qnbinom(0.975,size=8,
                                       prob=8/(8+likeCIdeath$Data))
likeCIdeath$Output_Var<-"deathIncrease"

likeCIpositive <- data.frame(Data=c(0.1,0.3,seq(1,
                                   max(subset(predobs.df,
                                              Output_Var=="positiveIncrease"
                                              )$Data)),20000))
likeCIpositive$Prediction.50. <- likeCIpositive$Data
likeCIpositive$Prediction.2.5. <- qnbinom(0.025,size=4,
                                       prob=4/(4+likeCIpositive$Data))
likeCIpositive$Prediction.97.5. <- qnbinom(0.975,size=4,
                                       prob=4/(4+likeCIpositive$Data))
likeCIpositive$Output_Var<-"positiveIncrease"
likeCI<-rbind(likeCIpositive,likeCIdeath)
likeCI$DataSet <- "Calibration"
likeCI.tmp <- likeCI
likeCI$DataSet <- "Validation"
likeCI <- rbind(likeCI, likeCI.tmp)

figS3<-ggplot(predobs.df)+
  geom_pointrange(aes(x=Data,y=Prediction.50.,
                      ymin=Prediction.2.5.,
                      ymax=Prediction.97.5.,
                      color=state,group=state),
                  position=position_jitter(width=0.1))+
  geom_ribbon(data=likeCI,
              aes(x=Data,ymin=Prediction.2.5.,
                      ymax=Prediction.97.5.,fill="CrI"),alpha=0.4)+
  geom_abline(slope=1,intercept=0)+
  scale_fill_viridis_d(option="magma",begin = 0.95,end=0.95)+
  scale_color_viridis_d()+
  coord_cartesian(xlim=c(0.1,20000),ylim=c(0.1,20000))+
  scale_x_log10()+
  scale_y_log10()+facet_grid(DataSet~Output_Var,labeller = labeller(Output_Var=obsnames))+
  ylab("Prediction (95% CrI)")+guides(fill = "none")+
  theme_bw()+
  annotation_logticks()
print(figS3)
ggsave("FigS3.pdf",plot=figS3,height=6,width=8,scale=1.5)
ggsave("FigS3.jpeg",plot=figS3,height=6,width=8,scale=1.5)
```

Each panel shows data versus predictions for daily deaths (left column) or cases (right column), with calibration results shown in the top panels (through April 1) and validation results in the bottom panels (through April 30).  Each point includes the 95% CrI for the prediction. The band centered on the y=x line shows the additional 95% CrI for the negative binomial likelihood, which is used to account for day-to-day variation in reporting.
