---
title: "Mobility Fits"
author: "Weihsueh Chiu"
date: "6/20/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height=7,fig.width=7,dpi = 300)
library(tidyverse)
library(covid19us)
library(jsonlite)
library(rvest)
library(data.table)
library(RCurl)
library(nlme)
library(minpack.lm)
library(ggpubr)
datezero<-"2019-12-31"
datemax<-as.numeric(as.Date("2020-06-03")) -as.numeric(as.Date(datezero))
fit0datemax<-as.numeric(as.Date("2020-04-10")) -as.numeric(as.Date(datezero))
datadir <- "../data"
get_fips2 <- function(datadir="../data") {
  fips_url<-"https://en.wikipedia.org/wiki/Federal_Information_Processing_Standard_state_code"
  fips_table<-fips_url%>%
    read_html()%>%
    html_nodes(xpath='//*[@id="mw-content-text"]/div/table[1]')%>%
    html_table%>%
    as.data.frame()%>%
    filter(Alpha.code!="")%>%
    mutate(Numeric.code=as.character(Numeric.code))%>%
    mutate(Numeric.code=str_pad(Numeric.code, width=2, side="left", pad="0"))
  fips_table$numFIPS <- as.numeric(fips_table$Numeric.code)
  fips_table <- subset(fips_table,numFIPS<=56)
  pop<-read.csv(file.path(datadir,"SCPRC-EST2019-18+POP-RES.csv"))
  pop <- subset(pop,STATE <=56)
  rownames(pop) <- as.character(pop$STATE)
  fips_table$pop <- pop[as.character(fips_table$numFIPS),"POPESTIMATE2019"]
  fips_table <- rbind(data.frame(Name="United States",
                                 Alpha.code="US",
                                 Numeric.code="00",
                                 Status="Country",
                                 numFIPS=0,
                                 pop=pop$POPESTIMATE2019[
                                   pop$STATE==0
                                   ]),
                      fips_table)
  rownames(fips_table) <- fips_table$Alpha.code
  return(fips_table)
}

get_udata_states <- function(fips_table,
                             datezero='2019-12-31',
                             datadir="../data") {
  fips_vec<-fips_table$Numeric.code
  row.names(fips_table)<-fips_table$Alpha.code
  fname<-file.path(datadir,paste0("UDATA-v3-full-state.",Sys.Date(),".csv"))
  if (file.exists(fname)) {
    udata<-read.csv(fname,as.is=TRUE,colClasses = c(rep("character",2),
                                                    rep("numeric",3)))
  } else {
    l<-list()
    for (i in 1:length(fips_vec)){
      l[[i]]<-fromJSON(paste0("https://covid19-scoreboard-api.unacastapis.com/api/search/covidstateaggregates_v3?q=stateFips:",fips_vec[i],"&size=10000"))
    }
    m<- Filter(function(x){length(x)>0 && x[["hits"]][["total"]][["value"]] > 0},l)
    unacast_list<-list()
    for (i in 1:length(m)){
      temp<-m[[i]]
      names(temp$hits$hits$`_source`$data)<-temp$hits$hits$`_source`$stateFips
      unacast_list[[i]]<-rbindlist(temp$hits$hits$`_source`$data,idcol = T)%>%
        select(.id,date,totalGrade,travelDistanceMetric,visitationMetric,encountersMetric)%>%
        group_by(.id)
    }
    unacast_df<-as.data.frame(rbindlist(unacast_list))
    udata <- unacast_df[,-3]
    write.csv(udata,fname,row.names=FALSE)
  }
  udata$date <- as.Date(udata$date)
  udata$numFIPS <- as.numeric(udata$.id)
  udata <- subset(udata,(numFIPS <= 56))
  rownames(fips_table) <- as.character(fips_table$Numeric.code)
  udata$State.Abbr <- fips_table[udata$.id,"Alpha.code"]
  udata$numDate <- as.numeric(as.Date(udata$date))-
    as.numeric(as.Date(datezero))
  udata$travelDistanceMetric<-udata$travelDistanceMetric+1
  udata$visitationMetric<-udata$visitationMetric+1
  udata$encountersMetric<-udata$encountersMetric+1
  names(udata)[names(udata)=="travelDistanceMetric"]<-"UnacastTravelDistance"
  names(udata)[names(udata)=="visitationMetric"]<-"UnacastVisitation"
  names(udata)[names(udata)=="encountersMetric"]<-"UnacastEncounters"
  udata.df <- melt(as.data.table(udata),
                   id.vars=c(".id","date","State.Abbr",
                             "numFIPS","numDate"),
                   measure.vars=c("UnacastTravelDistance",
                                  "UnacastVisitation",
                                  "UnacastEncounters"
                                  ))
  udata.df <- udata.df[order(udata.df$variable,udata.df$numDate,
                             udata.df$numFIPS),]
  return(udata.df)
}

func0 <- function(t,ys,thetamin,tautheta,ntheta) {
  s <- exp(-((t-60)/tautheta)^ntheta)
  y <- ys*(thetamin + (1-thetamin)*s)
  return(y)
}
func1 <- function(t,ys,thetamin,tausalt,taur,rmax) {
  tr <- 60 + tausalt
  utr <- 1 - 1/(1+exp(1*(t - tr)))
  utrtaur <- 1 - 1/(1+exp(1*(t - (tr+taur))))
  r <- (t - tr) * (rmax/taur) * (utr - utrtaur) + rmax*utrtaur
  y <- ys*(thetamin + (1-thetamin)*r)
  return(y)
}
  
func2 <- function(t,ys,thetamin,tautheta,ntheta,taus,taur,rmax) {
  s <- exp(-((t-60)/tautheta)^ntheta)
  tr <- 60 + tautheta + taus
  utr <- 1 - 1/(1+exp(2*(t - tr)))
  utrtaur <- 1 - 1/(1+exp(2*(t - (tr+taur))))
  r <- (t - tr) * (rmax/taur) * (utr - utrtaur) + rmax*utrtaur
  y <- ys*(thetamin + (1-thetamin)*(s + r))
  return(y)
}

func <- function(t,ys,thetamin,tautheta,ntheta,taus,taur,rmax) {
  s <- exp(-((t-60)/tautheta)^ntheta)
  tr <- 60 + tautheta + taus
  utr <- 1 - 1/(1+exp(4*(t - tr)))
  utrtaur <- 1 - 1/(1+exp(4*(t - (tr+taur))))
  r <- (t - tr) * (rmax/taur) * (utr - utrtaur) + rmax*utrtaur
  y <- ys*(thetamin + (1-thetamin)*(s + r))
  return(y)
}

mobfit <- function(mdata.df,pdfname="",showplots=TRUE,
                   minmaxinit = FALSE) {
  mfits.df <- subset(mdata.df,numDate >= 60)
  mfits.df$fit <- 0
  mfits.df$ys <- 0
  mfits.df$thetamin <- 0
  mfits.df$tautheta <- 0
  mfits.df$ntheta <- 0
  mfits.df$taus <- 0
  mfits.df$taur <- 0
  mfits.df$rmax <- 0
  statenow<-"DE"
  #varnow<-"travelDistanceMetric"
  for (statenow in unique(as.character(mfits.df$State.Abbr))) {
    for (varnow in unique(as.character(mfits.df$variable))) {
      dattmp<-subset(mfits.df,State.Abbr==statenow &
                             variable==varnow)
      dattmp$value <- runmed(dattmp$value,7,endrule="constant")
      dattmp0<-subset(dattmp,numDate <= fit0datemax)
      ys0<-mean(first(dattmp0$value,5))
      thetamin0<-mean(subset(dattmp0,date >= "2020-03-30" & 
                                   date <= "2020-04-03")$value)/ys0
      tautheta0<-20
      ntheta0<-10
      # plot(dattmp0$numDate,dattmp0$value)
      # lines(dattmp0$numDate,func0(dattmp0$numDate,
      #                            ys0,
      #                            thetamin0,
      #                            tautheta0,
      #                            ntheta0),lty=2)
      fittmp0<-nlsLM(value~func0(numDate,ys,thetamin,tautheta,ntheta),
               data=dattmp0,
               start = list(ys = ys0,
                            thetamin = thetamin0, 
                            tautheta = tautheta0,
                            ntheta = ntheta0))
      # lines(dattmp0$numDate,predict(fittmp0),lwd=2)
      ys0<-coef(fittmp0)["ys"]
      thetamin0<-coef(fittmp0)["thetamin"]
      tautheta0<-coef(fittmp0)["tautheta"]
      ntheta0<-coef(fittmp0)["ntheta"]
      #####
      dattmp1<-subset(dattmp,numDate > (fit0datemax-10))
      #thetamin1<-mean(mean(first(dattmp1$value,7)))/ys0
      thetamin1<-min(dattmp1$value)/ys0
      if (minmaxinit) {
        tausalt0 <- min(dattmp1$numDate[
          which(dattmp1$value > 
            (min(dattmp1$value) + 
               0.1*(max(dattmp1$value)-min(dattmp1$value))))]) - 60
        taur0 <- max(dattmp1$numDate[
          which(dattmp1$value < 
            (min(dattmp1$value) + 
               0.9*(max(dattmp1$value)-min(dattmp1$value))))]) - tausalt0 - 60
        taur0 <- max(7,taur0)

      } else {
        tausalt0<-fit0datemax+5-60
        taur0 <- max(7,last(dattmp1$numDate)-tausalt0-60-10)
      }
      rmax0<-((mean(last(dattmp1$value,5))/ys0)-thetamin1)/
        (1-thetamin1)
      # plot(dattmp1$numDate,dattmp1$value)
      # lines(dattmp1$numDate,func1(dattmp1$numDate,
      #                            ys0,
      #                            thetamin0,
      #                            tausalt0,
      #                            taur0,
      #                            rmax0),lty=2)
      if (is.character(try(
        fittmp1<-nlsLM(value~func1(numDate,ys0,thetamin,tausalt,taur,rmax),
               data=dattmp1,
               start = list(thetamin = thetamin1,
                            tausalt = tausalt0,
                            taur = taur0,
                            rmax = rmax0)),TRUE))) {
        print(paste("Error fit1",statenow,varnow))
        plot(dattmp1$numDate,dattmp1$value)
        lines(dattmp1$numDate,func1(dattmp1$numDate,
                                   ys0,
                                   thetamin1,
                                   tausalt0,
                                   taur0,
                                   rmax0),lty=2)
                            }
      # lines(dattmp1$numDate,predict(fittmp1))
      thetamin0 <- mean(thetamin0,coef(fittmp1)["thetamin.ys"])
      taus0<-coef(fittmp1)["tausalt"]-tautheta0
      taur0<-coef(fittmp1)["taur"]
      rmax0<-coef(fittmp1)["rmax.ys"]
      #taus0<- 28 # 5 fails
      #taur0<-max(7,last(dattmp$numDate)-tautheta0-taus0-60)
      # rmax0<-((mean(last(dattmp$value,5))/ys0)-thetamin0)/
      #   (1-thetamin0)
      # plot(dattmp$numDate,dattmp$value)
      # lines(dattmp$numDate,func(dattmp$numDate,ys0,thetamin0,tautheta0,ntheta0,
      #                           taus0,taur0,rmax0),lty=2)
      if(is.character(try({
        fittmp2<-nlsLM(value~func2(numDate,ys,thetamin,tautheta,ntheta,
                                 taus,taur,rmax),
               data=dattmp,
               start = list(ys = ys0,
                            thetamin = thetamin0, 
                            tautheta = tautheta0,
                            ntheta = ntheta0,
                            taus = taus0,
                            taur = taur0,
                            rmax = rmax0));
        fittmp<-nlsLM(value~func(numDate,ys,thetamin,tautheta,ntheta,
                                 taus,taur,rmax),
               data=dattmp,
               start = list(ys = coef(fittmp2)[1],
                            thetamin = coef(fittmp2)[2], 
                            tautheta = coef(fittmp2)[3],
                            ntheta = coef(fittmp2)[4],
                            taus = coef(fittmp2)[5],
                            taur = coef(fittmp2)[6],
                            rmax = coef(fittmp2)[7]));
      rowsnow<-mfits.df$State.Abbr==statenow &
                            mfits.df$variable==varnow;
      mfits.df[rowsnow,"fit"]<-predict(fittmp);#func(dattmp$numDate,
                                    #coef(fittmp)[1],
                                    #coef(fittmp)[2],
                                    #coef(fittmp)[3],
                                    #coef(fittmp)[4],
                                    #coef(fittmp)[5],
                                    #coef(fittmp)[6],
                                    #coef(fittmp)[7]
                                    #);#predict(fittmp);
      mfits.df[rowsnow,"ys"]<-coef(fittmp)[1];# ["ys"];
      mfits.df[rowsnow,"thetamin"]<-coef(fittmp)[2]; #["thetamin"];
      mfits.df[rowsnow,"tautheta"]<-coef(fittmp)[3]; #["tautheta"];
      mfits.df[rowsnow,"ntheta"]<-coef(fittmp)[4]; #["ntheta"];
      mfits.df[rowsnow,"taus"]<-coef(fittmp)[5]; #["taus"];
      mfits.df[rowsnow,"taur"]<-coef(fittmp)[6]; #["taur"];
      mfits.df[rowsnow,"rmax"]<-coef(fittmp)[7]; #["rmax"];
      }))) {
        print(paste("Error fit",statenow,varnow))
        plot(dattmp$numDate,dattmp$value)
        lines(dattmp$numDate,func(dattmp$numDate,ys0,thetamin0,tautheta0,ntheta0,
                                  taus0,taur0,rmax0),lty=2)
      }
    }
  }
  plist<-list()
  if (pdfname!="") {
    pdf(pdfname,height=8.5,width=11)
    for (varnow in unique(as.character(mfits.df$variable))) {
      p<-ggplot(subset(mdata.df,variable==varnow))+
               geom_point(aes(x=date,y=value))+
        geom_point(aes(x=as.Date(60+tautheta+taus,origin=datezero),
                       y=fit),color="red",
                   data=subset(mfits.df[mfits.df$numDate==
                                          round(60+mfits.df$tautheta+mfits.df$taus),],
                               variable==varnow))+
                 geom_line(aes(x=date,y=fit),color="red",
                           data=subset(mfits.df,variable==varnow))+
                 facet_wrap(~State.Abbr,scales="free")+
                 geom_vline(aes(xintercept=value,linetype=ReopenType,
                                color=ReopenType),
                            data=reopendat.df)+
                 theme(legend.position = "bottom")
      print(p+ggtitle(varnow))
      plist<-c(plist,list(p+ggtitle(varnow)))
    }
    dev.off()
  }

  if (showplots) {
    for (j in 1:length(plist)) {
      print(plist[[j]])
    }
  }
  return(mfits.df)
}

## Load FIPS
fipsname<-file.path(datadir,"FIPS_TABLE2.csv")
if (file.exists(fipsname)) {
  fips_table<-read.csv(fipsname,as.is=TRUE,row.names=1,
                       colClasses = c(rep("character",5),rep("numeric",2)))
} else {
  fips_table<-get_fips2()
  write.csv(fips_table,file=fipsname)
}

## Load re-opening times as documented by Washington Post.

reopendat <- read.csv(file.path(datadir,"ReopeningData.csv"),as.is=TRUE)
names(reopendat)[2]<-"State.Abbr"
reopendat[reopendat==""]<-NA
for (j in 4:10) reopendat[[j]]<-as.Date(reopendat[[j]])
reopendat.df <- melt(as.data.table(reopendat),id.vars=1:3)
names(reopendat.df)[4]<-"ReopenType"
reopendat.df$numDate <- as.numeric(reopendat.df$value)-
    as.numeric(as.Date(datezero))

```

## Mobility data

Each mobility dataset for each state is Fit to Weibull+linear model, where the Weibull represents the process of sheltering in place, and linear represents the re-opening phase.  The model is formulated as follows:

$$
y  = y_s\left\{\theta_{min}+(1-\theta_{min})\cdot\left[s(t) + r(t)\right]\right\} 
$$

$$
s(t)  =  \exp\left[-(t/\tau_\theta)^{n_\theta}\right]
$$

$$
r(t)  =  r_{max}\frac{t-\tau_\theta-\tau_s}{\tau_r}\left[
u(t-t_r)-
u(t-t_{rmax})\right] + u(t-t_{rmax})
$$
$$
u(t)  =  \text{Heaviside}(t)\approx 1 - \frac{1}{1+e^{4t}}
$$
$$
t_r  =  \tau_\theta+\tau_s
$$
$$
t_{rmax}  =  \tau_\theta+\tau_s+\tau_r
$$
  
Here $y_s$ represents the baseline (pre-pandemic) value, and the Weibull term has parameters $\theta_{min}$, $\tau_\theta$, and $n_\theta$.  The "reopening" term $r(t)$ is 0 prior to $t_r$, linear between $t_r$ and $t_{rmax}$, and constant at a value of $r_{max}$ after that, and made continuous by approximating the Heaviside function by a logistic function.  The reopening time is defined as $\tau_s$ days after $\tau_\theta$, and the maximum reopening amount $r_{max}$ happens $\tau_r$ days after that.  The time-dependence of both the sheltering phase $s(t)$ and reopening phase $r(t)$ are represented by functions with a range between 0 and 1.

### Unacast

```{r udata, echo=FALSE,fig.height=7,fig.width=7}

udata.df <- get_udata_states(fips_table)
ufits.df<- mobfit(udata.df,pdfname="Unacast.pdf")
```


### Google

```{r google, echo=FALSE,fig.height=7,fig.width=7}

fname<-file.path(datadir,paste0("google_mobility.",Sys.Date(),".csv"))
if (!file.exists(fname)) {
  download.file("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv",fname)
}
gdata <- read.csv(fname,as.is=TRUE)
gdata$date <- as.Date(gdata$date)
gdata <- subset(gdata,country_region_code=="US" &
                  sub_region_1!="" &
                  sub_region_2=="")
gdata$residential_neg_percent_change_from_baseline <- 
  -gdata$residential_percent_change_from_baseline
names(gdata)[names(gdata)=="retail_and_recreation_percent_change_from_baseline"]<-
  "Google_retail_and_recreation"
names(gdata)[names(gdata)=="grocery_and_pharmacy_percent_change_from_baseline"]<-
  "Google_grocery_and_pharmacy"
names(gdata)[names(gdata)=="parks_percent_change_from_baseline"]<-
  "Google_parks"
names(gdata)[names(gdata)=="transit_stations_percent_change_from_baseline"]<-
  "Google_transit_stations"
names(gdata)[names(gdata)=="workplaces_percent_change_from_baseline"]<-
  "Google_workplaces"
names(gdata)[names(gdata)=="residential_neg_percent_change_from_baseline"]<-
  "Google_residential_neg"
names(gdata)[names(gdata)=="residential_percent_change_from_baseline"]<-
  "Google_residential"

row.names(fips_table)<-as.character(fips_table$Name)
gdata$State.Abbr <- as.character(fips_table[gdata$sub_region_1,"Alpha.code"])
gdata$numDate <- as.numeric(as.Date(gdata$date))-
    as.numeric(as.Date(datezero))
gdata.df <- melt(as.data.table(
  gdata[,c("sub_region_1","State.Abbr","numDate","date",
           "Google_retail_and_recreation",
           #"Google_grocery_and_pharmacy",
           #"Google_parks",
           "Google_transit_stations",
           "Google_workplaces",
           "Google_residential_neg"
           )]),id.vars=1:4)
gdata.df$value <- (gdata.df$value+100)/100

# ggplot(gdata.df)+geom_line(aes(x=numDate,y=value,color=variable))+facet_wrap(~State.Abbr,scales="free")+theme(legend.position = "bottom")
gfits.df<-mobfit(gdata.df,pdfname="GoogleMobility.pdf")
```

### OpenTable

If missing, substitute US country-wide value.

```{r opentable, echo=FALSE,fig.height=7,fig.width=7}

fname<-file.path(datadir,paste0("YoY_Seated_Diner_Data.",Sys.Date(),".csv"))
n<-1
while (!file.exists(fname)) {
  fname<-file.path(datadir,paste0("YoY_Seated_Diner_Data.",Sys.Date()-n,".csv"))
  n<-n+1
}
odata<-read.csv(fname,as.is=TRUE)
odata<-subset(odata,Type=="state" | Name=="United States")
names(odata)<-gsub("X","2020-",names(odata))
names(odata)<-gsub("\\.","-",names(odata))
odata.df <- melt(as.data.table(odata),id.vars=1:2,stringsAsFactors=FALSE)
odata.df$variable<-as.Date(odata.df$variable)
names(odata.df)[names(odata.df)=="variable"]<-"date"
odata.df$numDate <- as.numeric(as.Date(odata.df$date))-
    as.numeric(as.Date(datezero))
rownames(fips_table)<-fips_table$Name
odata.df$numFIPS <- fips_table[as.character(odata.df$Name),"numFIPS"]
odata.df <- subset(odata.df,!is.na(numFIPS))
odata.df <- subset(odata.df,numFIPS <=56)


odata.df$State.Abbr <- fips_table[as.character(odata.df$Name),"Alpha.code"]
odata.df$value <- (odata.df$value+100)/100
odata.df$variable <- "OpenTable Bookings"

odata.us <- subset(odata.df,Name=="United States")
odata.df <- subset(odata.df,Name!="United States")
for (j in 2:nrow(fips_table)) {
  if (!(fips_table[j,"Name"] %in% odata.df$Name)) {
    odata.tmp <- odata.us
    odata.tmp$Type <- "state"
    odata.tmp$Name <- fips_table[j,"Name"]
    odata.tmp$numFIPS <- fips_table[j,"numFIPS"]
    odata.tmp$State.Abbr <- as.character(fips_table[j,"Alpha.code"])
    odata.df <- rbind(odata.df,odata.tmp)
  } 
}
# ggplot(odata.df)+geom_line(aes(x=numDate,y=value,color=variable))+facet_wrap(~State.Abbr,scales="free")+theme(legend.position = "bottom")

ofits.df<-mobfit(odata.df,pdfname="OpenTable.pdf",minmaxinit = TRUE)
```



## Summary by state

```{r by state}
vars <- c("date","State.Abbr","numDate","variable","value","fit",
              "ys","thetamin","tautheta","ntheta",
                                 "taus","taur","rmax")
allfits.df <- rbind(
  ufits.df[,..vars],
  gfits.df[,..vars],
  ofits.df[,..vars]
)
allfits.df$normfit <- allfits.df$fit/allfits.df$ys
allfits.df$scalefit <- (allfits.df$fit/allfits.df$ys - allfits.df$thetamin)/(1-allfits.df$thetamin)
allfits.df$treopen <- 60+allfits.df$tautheta+allfits.df$taus
names(allfits.df)[names(allfits.df)=="variable"]<-"MobilityMetric"
allfits.df$AbbrMobMet<-abbreviate(gsub("_","",
                                             allfits.df$MobilityMetric),
                                        minlength=10)
write.csv(allfits.df,file="MobilityFits.csv",row.names=FALSE)

allfitstmp.df<-allfits.df
allfitstmp.df$ys <- log10(allfitstmp.df$ys)
names(allfitstmp.df)[names(allfitstmp.df)=="ys"]<-"log10ys"
fitparms.df <- melt(as.data.table(
  allfitstmp.df[,c("date","State.Abbr","MobilityMetric","AbbrMobMet",
                "log10ys","thetamin","tautheta","ntheta",
                                 "taus","taur","rmax","treopen")]),
  id.vars=1:4)
fitparms.df <- subset(fitparms.df,date==as.Date(fit0datemax,origin=datezero))
write.csv(fitparms.df,file="MobilityParms.csv",row.names=FALSE)

pdf("Mobility.pdf",height=8.5,width=11)
for (statenow in unique(as.character(allfits.df$State.Abbr))) {
  p<-ggplot(subset(allfits.df,State.Abbr==statenow))+
           geom_point(aes(x=date,y=value))+
             geom_line(aes(x=date,y=fit),color="red",
                       data=subset(allfits.df,State.Abbr==statenow))+
             geom_vline(aes(xintercept=value,linetype=ReopenType,
                            color=ReopenType),
                        data=subset(reopendat.df,State.Abbr==statenow))+
      geom_point(aes(x=as.Date(treopen,origin=datezero),
                   y=fit),color="red",
               data=subset(allfits.df[allfits.df$numDate==
                                      round(allfits.df$treopen),],
                           State.Abbr==statenow))+
             facet_wrap(~MobilityMetric,scales="free")+
             theme(legend.position = "bottom")
  print(p+ggtitle(statenow))
  pp <- ggplot(subset(fitparms.df,State.Abbr==statenow))+
    geom_boxplot(aes(y=value),outlier.color="grey")+
    geom_point(aes(x=AbbrMobMet,y=value))+
    facet_wrap(~variable,scales="free")+coord_flip()
  print(pp)
}
dev.off()

pdf("MobilityFitNorm.pdf",height=6,width=9)
for (statenow in unique(as.character(allfits.df$State.Abbr))) {
  p<-ggplot(subset(allfits.df,State.Abbr==statenow))+
             geom_line(aes(x=date,y=normfit,color=MobilityMetric,
                           linetype=MobilityMetric),
                       data=subset(allfits.df,State.Abbr==statenow))+
      geom_point(aes(x=as.Date(treopen,origin=datezero),
                   y=normfit),color="red",
               data=subset(allfits.df[allfits.df$numDate==
                                      round(allfits.df$treopen),],
                           State.Abbr==statenow))+
             theme(legend.position = "bottom")
  print(p+ggtitle(statenow))
}
dev.off()

pdf("MobilityFitScale.pdf",height=6,width=9)
for (statenow in unique(as.character(allfits.df$State.Abbr))) {
  p<-ggplot(subset(allfits.df,State.Abbr==statenow))+
             geom_line(aes(x=date,y=scalefit,color=MobilityMetric,
                           linetype=MobilityMetric),
                       data=subset(allfits.df,State.Abbr==statenow))+
      geom_point(aes(x=as.Date(treopen,origin=datezero),
                   y=scalefit),color="red",
               data=subset(allfits.df[allfits.df$numDate==
                                      round(allfits.df$treopen),],
                           State.Abbr==statenow))+
             theme(legend.position = "bottom")
  print(p+ggtitle(statenow))
}
dev.off()
```

## Discussion

The Weibull+linear parameterization gives an adequate fit to the time-dependence for all the mobility data sources investigated.  Although clearly the linear increase cannot continue indefinitely, at least through the beginning of June, it is fairly representative.

Of note is that for all mobility data sources except for restaurant reservations, the linear increase began *prior* to the first re-opening milestone.  Therefore, it is not necessarily reliable to rely upon announced re-opening dates to represent when changes in transmission rates may occur.

```{r distributions of parameters}

fitm<-as.data.frame(aggregate(value~variable*State.Abbr,data=fitparms.df,mean))
names(fitm)<-c("variable","State.Abbr","value.mean")
fitsd<-as.data.frame(aggregate(value~variable*State.Abbr,data=fitparms.df,sd))
names(fitsd)<-c("variable","State.Abbr","value.sd")
fitq<-aggregate(value~variable*State.Abbr,data=fitparms.df,quantile,
          prob=c(0,0.025,0.25,0.5,0.75,0.975,1),simplify=TRUE)
fitsum<-cbind(fitm,fitsd,fitq)
fitsum<-fitsum[,-c(4:5,7:8)]
print(fitsum)
write.csv(fitsum,file="MobilityParmsSummaryByState.csv",row.names=FALSE)

ys0 <- median(allfits.df$ys)
thetamin0 <- median(allfits.df$thetamin)
tautheta0 <- median(allfits.df$tautheta)
ntheta0 <- median(allfits.df$ntheta)
taus0 <- median(allfits.df$taus)
taur0 <- median(allfits.df$taur)
rmax0 <- median(allfits.df$rmax)
treopen0 <- 60+tautheta0+taus0
treopen1 <- treopen0 + taur0

exfit.df <- data.frame(numDate=60:max(allfits.df$numDate))
exfit.df$date <- as.Date(exfit.df$numDate,origin=datezero)
exfit.df$y <- func(exfit.df$numDate,1,thetamin0,tautheta0,ntheta0,
                                 taus0,taur0,rmax0)
extau.df <- data.frame(numDate=60+c(tautheta0,
                             tautheta0+taus0,
                             tautheta0+taus0+taur0),
                       Milestone=factor(c("Shelter","ReopenStart","ReopenMax"),
                                        levels=c("Shelter","ReopenStart","ReopenMax"))
                       )
extau.df$date <- as.Date(extau.df$numDate,origin=datezero)
extau.df$y<- func(extau.df$numDate,1,thetamin0,tautheta0,ntheta0,
                                 taus0,taur0,rmax0)


pex<-ggplot(exfit.df)+
  geom_line(aes(x=date,y=y),size=2,color="grey")+
  geom_point(data=extau.df,aes(x=date,y=y,color=Milestone))+
  geom_label(data=extau.df[1,],aes(x=date,y=0,
                                   label=deparse(bquote(tau[theta]))),
             parse=TRUE,
             hjust=-0.1,vjust=-0.2)+
    geom_label(data=extau.df[2,],aes(x=date,y=0,
                                     label=deparse(bquote(tau[theta]+tau[s]))),
             parse=TRUE,
             hjust=-0.1,vjust=-0.2)+
    geom_label(data=extau.df[3,],aes(x=date,y=0,
                                     label=deparse(bquote(tau[theta]+tau[s]+tau[r]))),
             parse=TRUE,
             hjust=1.1,vjust=-0.2)+
  geom_segment(data=extau.df,aes(x=date,xend=date,y=y,yend=0,
                                 color=Milestone),linetype="dashed")+
  geom_segment(data=extau.df[-1,],
               aes(x=as.Date(60,origin=datezero),
                   xend=date,y=y,yend=y,color=Milestone),
               linetype="dashed")+
  geom_label(data=extau.df[2,],aes(x=date,y=y),
             label=deparse(bquote(theta[min])), 
             parse = TRUE,hjust=1.1,vjust=1.1)+
  geom_label(data=extau.df[3,],aes(x=date,y=y),label=
               deparse(bquote(theta[min]+r[max](1-theta[min]))),
             parse=TRUE,hjust=1,vjust=-0.1)+
  scale_x_date(date_minor_breaks="1 day",expand = c(0, 0))+
  scale_y_continuous(expand = c(0, 0),limits=c(0,1))+
  theme_bw()+theme(legend.position="none")+
  xlab("Date")+ylab("Normalized Mobility Metric")+
  scale_color_viridis_d(option="magma",end=0.6)

ggsave("MobilityFitExample.pdf",pex,height=4,width=6,scale=0.75)
```

```{r reopen dates}

fitreopen.df<-rbind(fitparms.df,
                    data.frame(date=reopendat.df$value,
                               State.Abbr=reopendat.df$State.Abbr,
                               MobilityMetric=reopendat.df$ReopenType,
                               AbbrMobMet=abbreviate(reopendat.df$ReopenType,
                                        minlength=10),
                               variable="ReopenDates",
                               value=reopendat.df$numDate))
fitreopen.df<-subset(fitreopen.df,!is.na(value))

```
