---
title: "Check Case Data Trend"
author: "Weihsueh Chiu"
date: "7/30/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,dpi = 600)
library(tidyverse)
library(covid19us)
library(jsonlite)
library(rvest)
library(data.table)
library(RCurl)
datezero <- "2019-12-31"
functiondir <- "../functions"
source(file.path(functiondir,"get_fips_function.R"))
source(file.path(functiondir,"get_testdata_functions.R"))

```

## Load case data with 7-day rolling mean (centered)


```{r load mean data,echo=FALSE}
fips_table <- fread("FIPS_TABLE.csv")

testdat.df <- get_testdata()
testdat.df <- subset(testdat.df,variable=="positiveIncrease" & 
                       state %in% fips_table$Alpha.code[2:52] &
                       numDate > (as.numeric(as.Date("2020-05-15")) - as.numeric(as.Date(datezero))))
mean7dat.df <- get_meantestdata()
mean7dat.df <- subset(mean7dat.df,variable=="meanpositiveIncrease" & 
                       state %in% fips_table$Alpha.code[2:52] &
                       numDate > (as.numeric(as.Date("2020-05-15")) - as.numeric(as.Date(datezero))))
# mean14dat.df <- get_meantestdata(nmean = 14)
# mean14dat.df <- subset(mean14dat.df,variable=="meanpositiveIncrease" & 
#                        state %in% fips_table$Alpha.code[2:52] &
#                        numDate > (as.numeric(as.Date("2020-06-20")) - as.numeric(as.Date(datezero))))

mean7max <- subset(mean7dat.df,numDate > (as.numeric(as.Date("2020-06-20")) - 
                                            as.numeric(as.Date(datezero)))) %>%
  group_by(state) %>%
  mutate(ismax = (max(value,na.rm=TRUE) == value))
mean7max <- as.data.frame(subset(mean7max, ismax))

stateorder <- mean7max$state[!duplicated(mean7max$state)]
testdat.df$state<-factor(testdat.df$state,levels=stateorder)
mean7dat.df$state<-factor(mean7dat.df$state,levels=stateorder)
mean7max$state<-factor(mean7max$state,levels=stateorder)
# mean14max <- mean14dat.df %>%
#   group_by(state) %>%
#   mutate(ismax = (max(value,na.rm=TRUE) == value))
# mean14max <- subset(mean14max, ismax)
```

## Peaks in the range of concern

First we are only worred about states with peaks after June 20 and before July 22.  So looking at the graph, that leaves the states up to Georgia (GA).

```{r peak before July 22, echo=FALSE, fig.width=10,fig.height=8}

p<-ggplot(aes(x=date,y=value),data=testdat.df)+
  geom_col(aes(col="Daily reported cases",linetype="July 06"),fill=NA)+
  geom_line(data=mean7dat.df,aes(col="Rolling 7-day mean"))+
  scale_color_viridis_d(begin=0.2,end=0.8)+
  facet_wrap(~state,scales="free_y")+
  geom_vline(aes(xintercept=as.Date("2020-07-06"),linetype="July 06"))+
  geom_vline(aes(xintercept=as.Date("2020-07-16"),linetype="July 16"))+
  geom_vline(aes(xintercept=as.Date("2020-07-22"),linetype="July 22"))+
  geom_point(data=mean7max,aes(color="Max of 7-day mean"),size=3)+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90))
print(p)

mean7max <- subset(mean7max, date < as.Date("2020-07-22"))
mean7dat.df <- subset(mean7dat.df, state %in% mean7max$state)
testdat.df <- subset(testdat.df,state %in% mean7max$state)

```

## Look for re-opened and then re-sheltered states

Next, we are only worried about states that have clearly re-opened and then re-sheltering, so that excludes ME, DE, VT, NY, and DC.

```{r peak, echo=FALSE, fig.width=10,fig.height=8}

noreopen <- c("ME","DE","VT","NY","DC")
mean7max <- subset(mean7max, !(state %in% noreopen))
mean7dat.df <- subset(mean7dat.df, !(state %in% noreopen))
testdat.df <- subset(testdat.df, !(state %in% noreopen))
p <- ggplot(aes(x=date,y=value),data=testdat.df)+
  geom_col(aes(col="Daily reported cases",linetype="July 06"),fill=NA)+
  geom_line(data=mean7dat.df,aes(col="Rolling 7-day mean"),size=2)+
  scale_color_viridis_d(begin=0.2,end=0.8)+
  facet_wrap(~state,scales="free_y")+
  geom_vline(aes(xintercept=as.Date("2020-07-06"),linetype="July 06"))+
  geom_vline(aes(xintercept=as.Date("2020-07-16"),linetype="July 16"))+
  geom_vline(aes(xintercept=as.Date("2020-07-22"),linetype="July 22"))+
  geom_point(data=mean7max,aes(color="Max of 7-day mean"),size=3)+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90))
print(p)#+coord_cartesian(xlim=as.Date(c("2020-07-01","2020-07-30"))))
# ggsave("Check Case Data Trend-July.pdf",p+coord_cartesian(xlim=as.Date(c("2020-07-01","2020-07-30"))),
#        dpi=300,height=8,width=10)
```

That leaves 15 states where there is a potential "plateau" or "decline" indicating a reversal of re-opening.

```{r print 15 states,echo=FALSE}
print(mean7max)
```

